Conversion of FFCS ICE fleet to EV: a data-driven approach

### 2020-11-17 Chapter2/2_1_intro.tex
fixed missing numbers

Car sharing refers to a model of car rental where customers rent a car for a short period of time, usually for a few hours or less. One of its most interesting systems is the so called \textit{Free-Floating Car Sharing (FFCS)} system. The peculiarity of this system is that customers can pick and drop the car wherever in a geo-fence area. The most famous company is car2go which is present in 25 cities and 8 different countries, both in Europe and North America. Moreover, other country-based providers exists like \textit{Enjoy} in Italy, operative in 6 cities.

@@

In total, \tool collected about XXX of trips in X cities for car2go, working from 2017 to 2018. Instead, considering Enjoy, I get about YYYY rides from YY cities from 2017 to 2020
\mc{rivedere dati su quanti bookings ci sono}. The source code of \tool for research purposes.\footnote{\url{github.com/MobilityPolito/}}

The reminder of this chapter is structured as follows: section~\ref{sec:2_3_data_acquisition} describes in details the raw data structure and the data flow from providers' API to the middle stage. Section~\ref{sec:2_4_data_normalization} describes how the software implements the data raw elaboration to trips records and their storage into data lake . \mc{continuare con altre sezione se ne aggiungo}

>>>>

Car sharing refers to a model of car rental where customers rent a car for a short period of time, usually for a few hours or less. One of its most interesting systems is the so called \textit{Free-Floating Car Sharing (FFCS)} system. The peculiarity of this system is that customers can pick and drop the car wherever in a geo-fence area. The most famous company is car2go which is present in 23 cities and 8 different countries, both in Europe and North America. Moreover, other country-based providers exists like \textit{Enjoy} in Italy, operative in 6 cities.

@@

In total, \tool collected about 27 million of trips in 23 cities for car2go, working from 2017 to 2018. Instead, considering Enjoy, I get about 6\,6 million rides from 6 cities from May 2017 to June 2019. The source code of \tool for research purposes.\footnote{\url{github.com/MobilityPolito/}}

The reminder of this chapter is structured as follows: section~\ref{sec:2_3_data_acquisition} describes in details the raw data structure and the data flow from providers' API to the middle stage. Section~\ref{sec:2_4_data_normalization} describes how the software implements the data raw elaboration to trips records and their storage into data lake. Section~\ref{sec:2_5_data_analysyes} introduces the final stage of data collection pipeline: a model able to perform several statistics on the collected data. Finally, section~\ref{sec:2_6_conclusions} concludes the chapter.

### 2020-11-17 Chapter2/2_3_methodology.tex
abstract added
fixed missing numbers

\begin{table}
	\setlength{\tabcolsep}{2.3pt}
	\centering
	\caption{Overview of car2go's data}
	\begin{tabular}{lcccc}
		\hline
		City &  City Size [$km^2$]\footnote{} & Population \footnotemark[\value{footnote}] &  Avg. Fleat & Bookings\\ 
		\hline
		\hline
		Columbus			  & 17 & 892k & 187 & 186k \\
		Florence 	  			& 34 & 372k & 220 & 333k \\
		Denver 		  			 & 36 & 727k & 312 & 348k \\
		Austin 		   			  & 31 & 964k & 315 & 377k \\
		Frankfurt 			   & 40 & 701k & 242 & 505k \\
		Toronto 				 & 49 & 3120k & 400 & 536k \\
		Amsterdam 			& 38 & 854k & 314 & 573k \\
		Montreal 				& 49 & 1704k & 429 & 606k \\
		New York City 	   & 77 & 8522k & 500 & 739k \\
		Turin 					    & 47 & 874k & 396 & 868k \\
		Munich 					 & 61 & 1464k & 478 & 916k \\
		Washington DC 	  & 64 & 705k & 563 & 919k \\
		Stuttgart 				 & 59 & 632k & 486 & 1001k \\
		Seattle 				   & 92 & 744k & 710 & 1134k \\
		Calgary 				  & 47 & 1239k & 552 & 1176k \\
		Rome 					  & 65 & 2837k & 582 & 1240k \\
		Rheinland 			   & 82 & 1688k & 648 & 1421k \\
		Vienna 					  & 72 & 1915k & 688 & 1702k \\
		Madrid 					 & 42 & 3233k & 424 & 2092k \\
		Milan 					   & 77 & 1396k & 776 & 2223k \\
		Hamburg 			  & 82 & 1833k & 812 & 2561k \\
		Vancouver 			 & 78 & 631k & 977 & 2701k \\
		Berlin 					  & 125 & 3769k & 1009 & 3091k \\
		%\\ \hline
		\hline
		\label{tab:2_3_datasummary}
	\end{tabular}
\end{table}
\footnotetext{\url{wikipedia.org}}
I let \tool scrape car2go's data  \mc{prendere le date}. In total it is possible to count about 27 million bookings spread in 23 cities. The table \ref{tab:2_3_datasummary} reports a brief resume of all the bookings, parkings present in the data lake.

\mc{devo prendere le date anche di enjoy}

\section{Data Analysis}
The third and final stage is the data analysis phase in which analytics modules query the MongoDB and obtain statistics. I rely on the Python programming language with Pandas and the GeoPandas libraries to deal with the data, the city zone definitions, provided by transport engineers as a shapefile, a popular geospatial vector data format, and the Geographical Information Systems (GIS) for the spatial analyses. I choose Python as it offer a large number of  libraries that easily interact with the different technologies like GIS, maps and MongoDB. In particular the usage of GeoPandas allow me to easily perform geographic analysis and split the city in many areas (or zones) of any possible shapes. I present more detailed characterization in chapter \ref{chap:3_charact}.

\section{Conclusion}
In this chapter I described the software pipeline, named \tool I used to harvest and store data from real FFCS providers.

>>>>

\begin{table}
	\setlength{\tabcolsep}{2.3pt}
	\centering
	\caption{Overview of car2go's data}
	\begin{tabular}{lcccc}
		\hline
		City &  City Size [$km^2$]\footnote{} & Population \footnotemark[\value{footnote}] &  Avg. Fleat & Bookings\\ 
		\hline
		\hline
		Columbus			  & 583 & 892k & 187 & 186k \\
		Florence 	  			& 102 & 372k & 220 & 333k \\
		Denver 		  			 & 401 & 727k & 312 & 348k \\
		Austin 		   			  & 704 & 964k & 315 & 377k \\
		Frankfurt 			   & 248 & 701k & 242 & 505k \\
		Toronto 				 & 630 & 3120k & 400 & 536k \\
		Amsterdam 			& 219 & 854k & 314 & 573k \\
		Montreal 				& 431 & 1704k & 429 & 606k \\
		New York City(Manhattan) 	   & 59 & 1629 & 500 & 739k \\
		Turin 					    & 130 & 874k & 396 & 868k \\
		Munich 					 & 310 & 1464k & 478 & 916k \\
		Washington DC 	  & 177 & 705k & 563 & 919k \\
		Stuttgart 				 & 207 & 632k & 486 & 1001k \\
		Seattle 				   & 217 & 744k & 710 & 1134k \\
		Calgary 				  & 825 & 1239k & 552 & 1176k \\
		Rome 					  & 1287 & 2837k & 582 & 1240k \\
		Rheinland 			   &  - & 1688k & 648 & 1421k \\
		Vienna 					  & 414 & 1915k & 688 & 1702k \\
		Madrid 					 & 604 & 3233k & 424 & 2092k \\
		Milan 					   & 181 & 1396k & 776 & 2223k \\
		Hamburg 			  & 755 & 1833k & 812 & 2561k \\
		Vancouver 			 & 115 & 631k & 977 & 2701k \\
		Berlin 					  & 891 & 3769k & 1009 & 3091k \\
		%\\ \hline
		\hline
		\label{tab:2_3_datasummary}
	\end{tabular}
\end{table}

\begin{table}
	\setlength{\tabcolsep}{2.3pt}
	\centering
	\caption{Overview of Enjoy's data}
	\begin{tabular}{lcccc}
		\hline
		City &  City Size [$km^2$]\footnotemark[\value{footnote}] & Population \footnotemark[\value{footnote}] &  Avg. Fleat & Bookings\\ 
		\hline
		\hline
		Bologna			         &140    & 390k & 72 & 47k \\
		Catania			          &182    & 311k & 78 & 205k \\	
		Florence 	  			 & 102  & 372k & 82 & 279k \\
		Turin 					   & 130   & 874k & 251 & 983k \\
		Rome 					  & 1287 & 2837k & 601 & 2031k \\
		Milan 					   & 181   & 1396k & 755 & 3107k \\

		%\\ \hline
		\hline
		\label{tab:2_3_datasummary_enjoy}
	\end{tabular}
\end{table}

\footnotetext{\url{wikipedia.org}\label{fn:wiki}}
I let \tool scrape car2go's data from December 2016 to January 2018. In total it is possible to count about 27 million bookings spread in 23 cities. The same scraping procedure run for Enjoy as well, from May 2017 to June 2019.  The table \ref{tab:2_3_datasummary} reports a brief resume of all the car2g bookings, present in the data lake, while table \ref{tab:2_3_datasummary_enjoy} reports the same information for Enjoy.

\section{Data Analysis}
\label{sec:2_5_data_analysyes}
The third and final stage is the data analysis phase in which analytics modules query the MongoDB and obtain statistics. I rely on the Python programming language with Pandas and the GeoPandas libraries to deal with the data, the city zone definitions, provided by transport engineers as a shapefile, a popular geospatial vector data format, and the Geographical Information Systems (GIS) for the spatial analyses. I choose Python as it offer a large number of  libraries that easily interact with the different technologies like GIS, maps and MongoDB. In particular the usage of GeoPandas allow me to easily perform geographic analysis and split the city in many areas (or zones) of any possible shapes. I present more detailed characterization in chapter \ref{chap:3_charact}.

\section{Conclusions}
\label{sec:2_6_conclusions}
In this chapter I described the software pipeline, named \tool I used to harvest and store data from real FFCS providers.

### 2020-11-17 thesis.tex
fixed missing numbers

In the first part of my thesis, I described the software I developed in order to scrape from the web real combustion engine FFCS, from two providers: car2go and Enjoy. The data collection car2go data collection lasted from \mc{december 2016 to january 2018, collecting more than 28 million users' bookings spread in 25 cities} while the Enjoy data collection phase started \mc{may 2018 and lasted until BOH}, collectin about \mc{BOH} bookings in 6 cities.

>>>>

In the first part of my thesis, I described the software I developed in order to scrape from the web real combustion engine FFCS, from two providers: car2go and Enjoy. The data collection car2go data collection lasted from \mc{december 2016 to january 2018, collecting more than 27 million users' bookings spread in 23 cities} while the Enjoy data collection phase started in May 2017 and lasted until June 2019, recording about 6 million bookings in 6 cities.
